---
title: "Data Science Homework 1"
output: github_document

---
Hemangi Modi (UNI: hbm2126)


2022-09-22

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading dataset, include=FALSE}
data("penguins", package = "palmerpenguins")
```

```{r loading libraries, include=FALSE}
library(tidyverse)
```

# Problem 1

### Describing the "penguins" dataset

Below are some of the characteristics of the "penguins" dataset:

* The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns

* There are 8 variables in this dataset: the species of the penguin, the island the species is from, the length of the bill in millimeters, the depth of the bill in millimeters, the length of the flipper in millimeters, the body mass of the penguin in grams, the sex of the penguin, and the year

* The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` millimeters

### Scatterplot

```{r creating a scatterplot}
ggplot(penguins, aes(x= bill_length_mm, y = flipper_length_mm, col = species)) + geom_point()
```

```{r saving the scatterplot}
ggsave("penguins_Scatterplot.png")
```

# Problem 2

```{r creating a data frame}
hw1_df = tibble(
    random_samp = rnorm (10),
    vec_logical = random_samp > 0,
    vec_char = c("This", "is", "a", "character", "vector","that","is","ten","characters","long"),
    vec_factor = factor(c("Near","Far","Middle","Middle","Near","Far","Near","Middle","Far","Near"))
  )
```

```{r means of variables}
mean(pull(hw1_df,random_samp))
mean(pull(hw1_df,vec_logical))
mean(pull(hw1_df,vec_char))
mean(pull(hw1_df,vec_factor))

```

It is only possible to take the mean of the variables in the random sample from a standard Normal distribution and the logical vector. Taking the mean of the character and factor vectors returns 'NA'

```{r converting variables to numeric values, eval=FALSE}
as.numeric(pull(hw1_df,vec_logical))
as.numeric(pull(hw1_df,vec_char))
as.numeric(pull(hw1_df,vec_factor))
```

Using the as.numeric function on the logical vector returns zeroes and ones. For the character vector, the as.numeric function returns a set of 'NAs.' This explains why taking the mean of the character vector returns 'NA.' On the factor vector, the as.numeric assigns '3,' '2', or '1' to the different 'levels' within the factor vector. As a result, after using as.numeric on the factor vector, it is possible to take the mean but may not be meaningful as data stored in factor vectors generally tend to be categorical and there is no relative position/rank. 
