---
title: "Data Science Homework 1"
output: github_document

---
Hemangi Modi (UNI: hbm2126)


2022-09-22

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading dataset}
data("penguins", package = "palmerpenguins")
```

```{r loading libraries, include=FALSE}
library(tidyverse)
```

# Problem 1

### Describing the "penguins" dataset

Below are some of the characteristics of the "penguins" dataset:

* The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns

* There are 8 variables in this dataset: the species of the penguin, the island the species is from, the length of the bill in millimeters, the depth of the bill in millimeters, the length of the flipper in millimeters, the body mass of the penguin in grams, the sex of the penguin, and the year

* The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` millimeters

### Scatterplot

```{r creating a scatterplot}
ggplot(penguins, aes(x= bill_length_mm, y = flipper_length_mm, col = species)) + geom_point()
```

```{r saving the scatterplot}
ggsave("penguins_Scatterplot.png")
```

# Problem 2

```{r creating a data frame}
hw1_df = tibble(
    random_samp = rnorm (10),
    vec_logical = random_samp > 0,
    vec_char = c("This", "is", "a", "character", "vector","that","is","ten","characters","long"),
    vec_factor = factor(c("Near","Far","Middle","Middle","Near","Far","Near","Middle","Far","Near"))
  )
```

```{r means of variables}
mean(pull(hw1_df,random_samp))
mean(pull(hw1_df,vec_logical))
mean(pull(hw1_df,vec_char))
mean(pull(hw1_df,vec_factor))

```

```{r converting variables to numeric values, eval=FALSE}
as.numeric(pull(hw1_df,vec_logical))
as.numeric(pull(hw1_df,vec_char))
as.numeric(pull(hw1_df,vec_factor))
```

